NMRlipids III: Lipid--Cholesterol Interactions in Atomistic Molecular Dynamics Simulations

### 2023-03-24 MANUSCRIPT/manuscript.tex
Update text

\documentclass[aps,prl,superscriptaddress]{revtex4-2}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{booktabs}
\usepackage{xr}

@@

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}

\externaldocument{MANUSCRIPT/manuscriptSUPPL}

@@

\author{Fernando Favela-Rosales}
\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
%
\author{Peter Heftberger}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
%
\author{Matti Javanainen}
\email{matti.javanainen@helsinki.fi}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}
%
\author{Jesper J. Madsen}
\affiliation{Department of Global Health, College of Public Health}
\affiliation{University of South Florida}
%
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
%
\author{Markus Miettinen}
\affiliation{MPI}
%
\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}
%
\author{Georg Pabst}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
\affiliation{BioTechMed-Graz, Graz 8010, Austria}
%
\author{Thomas Piggot}
\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
\affiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}

@@

\date{\today}

@@

\todo{Complete scattering methods}

%Such scattering experiments are challenging to interpret for multi-component lipid bilayers \cite{pan12,Heftberger15,Marquardt15}, yet we still extract  thus area per molecule values, one of the main quantity used to compare simulations to experiments, have not been available for lipid cholesterol mixtures.

We measured small angle X-ray scattering (SAXS) data on multilamellar vesicles (MLVs) composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments. The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtain both contributions. In this work, only the form factors are discussed. The electron density profiles have been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, \textit{i.e.} volume distribution functions are modelled by individual Gaussians or error functions. CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}. Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. 
%Additional figures show the volume distribution functions and the resulting electron density profiles.
%Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)

@@

\paragraph{More and Less Ordered Populations:} The angles ($\theta$) between the C--H vectors of the last six carbons of the palmitate chain were extracted and the data across all CHOL concentrations was accumulated. These data were fit with two Gaussians centered at 0,
%
\begin{align}
    P(\theta)=a_\mathrm{less}\exp\left(-\frac{\theta}{2\sigma_\mathrm{less}}\right)+a_\mathrm{more}\exp\left(-\frac{\theta}{2\sigma_\mathrm{more}}\right),
    \label{eq:twogauss}
\end{align}
%
where the two standard deviations ($\sigma$) correspond to the more and less ordered components. After these $\sigma$ values were established for a given force field, the data at all CHOL concentrations was independently fitted by Eq.~\eqref{eq:twogauss} with fixed $\sigma$ values. These fits provided the pre-factors, whose ratio $a_\mathrm{more}/(a_\mathrm{more}+a_\mathrm{less})$ corresponds to the fraction of the more ordered component in the system.

This two-Gaussian approach assumes a ``two-population'' scheme, where the more and less ordered populations are present (at non-zero CHOL concentration), and that the addition of CHOL only affects their ratio, but not the nature. In an alternative picture of a ``collective transition'', there would be a single Gaussian, whose standard deviation $\sigma$ would vary systematically as a function of CHOL concentration.

\paragraph{Quantitative Quality Measure for Lipid Bilayer Structure:} The deviations (in \%) from experimental values across CHOL concentrations were used to quantify the quality of the lipid force fields. The density profiles (as a function of distance from membrane core along the membrane normal), oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains), and diffusion coefficients were interpolated to create continuous 2D matrix for density profiles and order parameters, and a continuous 1D vector for the diffusion coefficients. For the 2D matrices, the absolute values of the differences of the matrices from simulations and experiments were calculated and averaged over one dimension which was either the distance from membrane core or the carbon atom number, respectively. This resulted in a 1D deviation vector as a function of CHOL concentration. For diffusion coefficients, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration. All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \%.

@@

\todo{Results: volmaps of chains around cholesterol?}\\
\todo{Results: SI: RDFs of POPC/CHOL around CHOL}\\
\todo{FF minima analysis}

@@

Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16,??}. The form factor can be translated to an electron density profile, and area per lipid and bilayer thickness values can be extracted using the scattering density profile (SDP) model or its combination with MD simulations  \cite{pan12,Heftberger15,Marquardt15,??,doktorova2020molecular}. Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.

Thus, we compare here the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species. For reference, all form factor profiles from experiment and simulation are shown in Fig.~\ref{SIfig:scattering}. As the absolute and relative lobe amplitudes depend on simulation box size (Fig.~\ref{SIfig:scatteringvssize}), the minima in the form factor profiles provide a robust comparison between simulation and experiment. Thus, we have also traced the minima in these profiles, and they are plotted in Fig.~\ref{SIfig:ffminima}. As simulations and experiments were performed at somewhat different CHOL concentrations, we generated two-dimensional maps for the density profiles using interpolation (see Methods). These maps are shown in Fig.~\ref{fig:densmaps} for both simulations (top row) and experiment (bottom row, left column). The original electron density profiles used in the generation of these maps are shown in Fig.~\ref{SIfig:densprofs}.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\linewidth]{../FIGS/densitymaps.pdf}
  \caption{\label{fig:densmaps}%
  \textbf{Electron density profiles, thickness and area per lipid as a function of CHOL concentration.}
  %
  Top: Electron density maps for the simulations using four different force fields.
  %
  Bottom left: Electron density map for the experiment. The scaling is uniform between all simulations and experiment. The original electron density profiles are shown in Fig.~\ref{SIfig:densprofs}. The scattering experiments are interpolated to 298~K from Refs.~\citenum{pabst2000structural} (58.8~\AA{}$^2$), \citenum{kuvcerka2006structure} (67.5~\AA{}$^2$), and \citenum{Kucerka11} (63.5~\AA{}$^2$). The NMR data is measured in Ref.~\citenum{Leftin14} using either $^{13}$C (58.9~\AA{}$^2$) or $^2$H (58.3~\AA{}$^2$) NMR. The effect of system size on the density profiles in simulations is demonstrated in Fig.~\ref{SIfig:densprofssize} in the SI.
  %
  Bottom center: Bilayer thickness. Thickness is defined as twice the distance from the peak in electron density to the membrane core. Experimental data are extracted in a similar manner from electron density profiles obtained with X-ray scattering.
  %
  Bottom right: area per lipid measured by dividing the total membrane area by the number of phospholipids. 
  %
  The size-dependency of area per lipid is shown in Fig.~\ref{SIfig:aplvssize}.
  }
\end{figure}

Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region rich in electron-rich phoshorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase. However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields. CHOL renders the membrane thicker, resulting in the deviation of the high-density bands with increasing amount of CHOL. In CHARMM36, Slipids, and MacRog, there is clear saturation of this effect at around 30~mol-\% CHOL. In Lipid17, however, the membrane gets thicker until the maximum studied CHOL concentration of 47~mol-\%. In experiment, the effect of CHOL is generally weaker, and some saturation is evident at high CHOL concentrations. These trends are visible in the bottom middle panel in Fig.~\ref{fig:densmaps}, which shows the effect of CHOL on membrane thickness. Fig.~\ref{fig:densmaps} also demonstrates that CHARMM36 has the sharpest low- and high-density bands indicating, and the same is true for MacRog at higher cholesterol concentrations. This uniform membrane structures could indicate small undulations, whereas the less sharp Lipid17 and especially Slipids profiles could indicate a more flexible membrane.

Comparison against experiments (bottom left panel in Fig.~\ref{fig:densmaps}) indicates that the membrane structure is more uniform in the experiment, as the bands are sharper. This is also evident in the original electron density profiles in Fig.~\ref{SIfig:densprofs}. System size plays a role here, as demonstrated by the density maps calculated for the CHARMM36 simulations in three sizes, shown in Fig.~\ref{SIfig:densprofssize}. The larger the system, the larger the bilayer undulations are, which can smear the features of the density profiles. Still, even in the smallest system, the band intensities are not as localized as in the experiment.

@@

The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}. The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL). However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, \textit{i.e.} a certain amount of CHOL could be added to a phospholipid bilayer without increasing or decreasing its total area \cite{javanainen2017two}. We thus compared the behavior of the area per (POPC) lipid (APL) as a function of CHOL concentration in the bottom right panel of Fig.~\ref{fig:densmaps}. The dependence of APL on CHOL concentration follows the trends in thickness inversely in general, yet provide curious differences between the force fields. Lipid17 has the largest APL across the entire CHOL contration range, and the addition of CHOL always leads to an increase in the membrane area. MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration indicating a particularly strong condensation effect. The profiles for Slipids and CHARMM36 are very similar, and the partial area of CHOL is small or even zero until a concentration of 20~mol-\%. Importantly, the observed differences in CHOL condensing effect take place at the physiologically relevant CHOL concentration range \cite{van2008membrane}. Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained by \citeauthor{pabst2000structural}\cite{pabst2000structural} and \citeauthor{kuvcerka2006structure}\cite{kuvcerka2006structure} for POPC at 303~K.

@@

Order parameters for C--H bond vectors measured with with $^{13}$C or $^{2}$H NMR techniques, provide indirect information about the structural sampling of indidual molecules \cite{ollila16}. Moreover, these parameters are readily extracted from the simulation. An increase in acyl chain order parameters with added cholesterol in simulations and experiments can be explained by increased anti conformations in acyl chains \cite{ferreira13,??}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour.

Here, as the NMR experiment \cite{ferreira13} was again performed at different CHOL concentrations than simulations (and X-ray scattering measurements), we generated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments, and subtracted these maps to obtain the deviation of simulation from experiment. These maps are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{fig:SIfig:oleate}.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\linewidth]{../FIGS/OP_chains.pdf}
  \caption{\label{fig:OPmaps}%
  POPC acyl chain order parameter deviation from experiments. Data are shown for palmitate (top row) and oleate (bottom row) and for our force fields (columns). Negative values indicate that order is too high ($-S_\mathrm{CD}$ values too negative) in the simulations. The values that are within the estimated error range of 0.02 from experiment are considered to be correct ``correct'' and coloured in white. The order parameters calculated for the two or three identical carbons were very similar and thus averaged over in simulations, as they have identical values in experiments. For the C2 carbon of the oleate chain, the order parameters are forked in simulations and experiments. We thus calculated the differences for both the larger and smaller values, and averaged this deviation for the C2 column of oleate.
  }
\end{figure}

These original maps in Fig.~\ref{SIfig:orderparameters} demonstrate that the addition of cholesterol substantially increases the acyl chain C--H order parameters for all simulation models and experiments, which can be explained by increased anti conformations in acyl chains \cite{ferreira13,??}. The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13}
exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because
phase separation is not observed for this mixture \cite{ionova12,ferreira13}.

The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive way for a quantitative comparison of the behaviors of the force fields. Here, deviations in the range [$0.02$,0.02] are considered to fall within the error estimate and are thus shown in white. Blue indicates that the order parameters are too low in the simulation (their absolute value is larger, \textit{i.e.} the chains are too ordered). Overall, the simulation models behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.

@@

The head group order parameters are not largely affected by CHOL concentration in experiments \cite{ferreira13}, which has been discussed in our earlier work \cite{botan15}. The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL\cite{botan15} largely persist across all studied CHOL concentrations. Moreover, it has now become standard to consider them in the parameterization and validation of lipid force fields \cite{dickson2022lipid21,grote2020optimization,yu2021charmm36,yu2021semi}.

\todo{Also experimental cholesterol order parameters are available. Maybe these should be calculated from simulations as well.}

@@

The previous sections highlight that the agreement of the studied force fields with experiment depends on both the parameter in question or the CHOL concentration. To capture these features in an easily digestible form, we have calculated the relative deviations of the density profiles, the order parameters of the two acyl chains, and the diffusion coefficients between the experiments and all the force fields. The interpolated data were used, and the relative deviation from experiments was obtained by dividing the difference of simulated and experimental data by the latter. These data are shown in the two leftmost columns in Fig.~\ref{fig:deviation} as stacked plots.

>>>>

\documentclass[journal=jctcce]{achemso}
\usepackage[version=3]{mhchem} 
\usepackage[T1]{fontenc}       
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{xr}
\usepackage{multirow}
% \usepackage{rotating} 
% \usepackage{times}
% \usepackage{graphicx}
% \usepackage{setspace}
% \usepackage{amsmath}
% \usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}

@@

\newcommand*{\myexternaldocument}[1]{%
    \externaldocument{#1}%
    \addFileDependency{#1.tex}%
    \addFileDependency{#1.aux}%
}
\myexternaldocument{manuscriptSUPPL}

\author{Fernando Favela-Rosales}
\affiliation{Departamento de F\'isica, Centro de Investigaci\'on y de Estudios Avanzados del IPN, Apartado Postal 14-740, 07000 M\'exico D.F., M\'exico}
%
\author{Peter Heftberger}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
%
\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
\email{matti.javanainen@helsinki.fi}
%
\author{Jesper J. Madsen}
\affiliation{Department of Global Health, College of Public Health}
\alsoaffiliation{University of South Florida}
%
\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
%
\author{Markus Miettinen}
\affiliation{MPI}
%
\author{O. H. Samuli Ollila}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\alsoaffiliation{Institute of Biotechnology, University of Helsinki}
\email{samuli.ollila@helsinki.fi}
%
\author{Georg Pabst}
\affiliation{Institute of Molecular Biosciences, Biophysics Division, NAWI Graz, University of Graz, Graz 8010, Austria}
\alsoaffiliation{BioTechMed-Graz, Graz 8010, Austria}
%
\author{Thomas Piggot}
\affiliation{School of Chemistry, University of Southamptaon, Southampton SO17 1BJ, United Kingdom}
\alsoaffiliation{Chemical Biological and Radiological Sciences, Defence Science and Technology Laboratory, Porton Down, Salisbury, Wiltshire SP4 0JQ, United Kingdom}

@@

@@

@@

We measured small angle X-ray scattering (SAXS) data on multilamellar vesicles (MLVs) composed of POPC and CHOL with the latter present at 0--50~mol-\% with 5~mol-\% increments. The data were obtained at the EMBL BioSAXS beamline (Hamburg) using 20~keV photons at $T=300$~K and analyzed in terms of the SDP-GAP model described in Refs.~\citenum{Heftberger.2014} and \citenum{Heftberger15}. All lipids were purchased from Avanti Polar Lipids  (Alabaster, AL, USA) as dry powders and were used without any further purification. All other chemicals were obtained in pro analysis quality from Lactan (Graz, Austria). Fully hydrated MLVs were prepared using rapid solvent exchange as detailed  previously \cite{rieder2015optimizing,belivcka2017high}. This avoids the precipitation of cholesterol crystallites at high concentration \cite{buboltz1999novel} yielding non-phase separated samples up to 50~mol-\% cholesterol content.

The data from MLVs contain the structure factor (the crystalline lattice) and form factor in a convoluted fashion, yet by fitting the scattered intensity data we obtain both contributions. In this work, only the form factors are discussed. The electron density profiles have been modelled in terms of the SDP model \cite{heberle12,Kucerka08a,kucerka12}, \textit{i.e.} volume distribution functions are modelled by individual Gaussians or error functions. CHOL is particularly challenging to describe within the SDP model, and therefore form factors are often compared between simulation and experiment \cite{Kucerka08b} instead of density profiles, or CHOL density is modeled together with that of CH$_2$ groups \cite{pan12}. Here, cholesterol is accounted for by two Gaussians, as proposed by Jianjun Pan (private communication), and used in Refs.~\citenum{belivcka2017high} and \citenum{heftbergerthesis}. 
%Additional figures show the volume distribution functions and the resulting electron density profiles.
%Authors to consult and potentially include in publications using this data: Peter Heftberger (peter.heftberger@gmx.at), Georg Pabst (georg.pabst@uni-graz.at)

@@

\paragraph{More and Less Ordered Populations:} The angles ($\theta$) between the C--H vectors of the last six carbons of the palmitate chain were extracted and the data across all CHOL concentrations was accumulated. These data were fit with two Gaussians centered at 0,
%
\begin{align}
    P(\theta)=a_\mathrm{less}\exp\left(-\frac{\theta}{2\sigma_\mathrm{less}}\right)+a_\mathrm{more}\exp\left(-\frac{\theta}{2\sigma_\mathrm{more}}\right),
    \label{eq:twogauss}
\end{align}
%
where the two standard deviations ($\sigma$) correspond to the more and less ordered components. After these $\sigma$ values were established for a given force field, the data at all CHOL concentrations were independently fitted by Eq.~\eqref{eq:twogauss} with the fixed $\sigma$ values. These fits provided the pre-factors, whose ratio $a_\mathrm{more}/(a_\mathrm{more}+a_\mathrm{less})$ corresponds to the fraction of the more ordered component in the system.

This two-Gaussian approach assumes a ``two-population'' scheme, where the more and less ordered populations are present (at non-zero CHOL concentration), and that the addition of CHOL only affects their ratio, but not their natures. In an alternative picture of a ``collective transition'', there would be a single Gaussian, whose standard deviation $\sigma$ would decrease as a function of CHOL concentration as the membrane gets more ordered.

\paragraph{Quantitative Quality Measure for Lipid Bilayer Structure:} The deviations (in \%) from experimental values across CHOL concentrations were used to quantify the quality of the lipid force fields. 
%The density profiles (as a function of distance from membrane core along the membrane normal), 
The 1\textsuperscript{st} and 2\textsuperscript{nd} form factor minima locations, the electron density profiles (as a function of normal distance from bilayer center), the oleate and palmitate chain order parameters (as a function of carbon atoms in the acyl chains) and diffusion coefficients were interpolated to create continuous 2D matrices for the order parameters and density profiles, and a continuous 1D vector for the diffusion coefficients and form factor minima. For the 2D matrices, the absolute values of the differences of the matrices from simulations and experiments were calculated and averaged over the carbon atoms in the acyl chain (order parameters) or across the simulation system (density profiles). This resulted in 1D deviation vectors as a function of CHOL concentration. For diffusion coefficients and form factor minima, the absolute value of the difference of the interpolated 1D vectors of simulation and experimental data was calculated to provide deviation as a function of CHOL concentration. All these 1D vectors were normalized by dividing them by the experimental values to provide relative deviations in \%.

@@

\todo{Results: volmaps of chains around cholesterol?}\\
\todo{Results: SI: RDFs of POPC chains and CHOL around CHOL}\\
\todo{Unify colours in figures}\\

@@

Lipid bilayer dimensions can be experimentally accessed by measuring X-ray scattering form factor, which is related to the electron density along membrane normal \textit{via} a Fourier transform \cite{pan12,Heftberger15,Marquardt15,ollila16,??}. The form factor can be translated to an electron density profile, and area per lipid and bilayer thickness values can be extracted using the scattering density profile (SDP) model or its combination with MD simulations  \cite{Kucerka08a,pan12,Heftberger15,Marquardt15,??,doktorova2020molecular}. Sophisticated approaches for single component lipid bilayers are available but the data for multi-component systems are more difficult to interpret, and the used approaches involve severe approximations \cite{pan12,Heftberger15,Marquardt15,??}.

Thus, we compare here the form factors and the total electron density profiles which do not require the de-convolution of the electron density profiles into densities of the different molecular species or their segments. For reference, all scattering intensitities from experiment are shown in Figs.~\ref{SIfig:scattering1} and \ref{SIfig:scattering2}, the form factor profiles from experiment and simulation in Fig.~\ref{SIfig:scattering} and the density profiles from experiment and simulation in Fig.~\ref{SIfig:densprofs}.

As the absolute and relative lobe amplitudes depend on simulation box size (Fig.~\ref{SIfig:scatteringvssize}), the minima in the form factor profiles provide a robust comparison between simulation and experiment. Thus, we first traced the first two minima in the form factors, and they are plotted in Fig.~\ref{SIfig:ffminima}. As simulations and experiments were performed at somewhat different CHOL concentrations, we  interpolated the minima to the entire studied range of CHOL concentrations. The plots in Fig.~\ref{SIfig:ffminima} highlight that the addition of CHOL first shifts the first minima to smaller wave vector values in the experiment, which is reasonably well captured by the simulation force fields. CHARMM36 seems to be off more than the other three force fields. Above $\sim$25~mol-\% of CHOL, the location of the minimum shifts to larger wave vector values in the experiment, which is curiously not captured by any of the force fields. The experiment shows a steady shift of the second minimum to smaller wave vector values, and this is reproduced by all simulation force fields. Slipids and Lipid17 are generally in best agreement with the experiment, whereas MacRog and CHARMM36 show poorer agreement.

Moving on to the electron density profiles extracted from the form factors, we generated two-dimensional maps for the density profiles using interpolation of the density profiles in Fig.~\ref{SIfig:densprofs} (see Methods). These maps are shown in Fig.~\ref{fig:densmaps} for both simulations (top row) and experiment (bottom row, left column).

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\linewidth]{../FIGS/densitymaps.pdf}
  \caption{\label{fig:densmaps}%
  \textbf{Electron density profiles, thickness and area per lipid as a function of CHOL concentration.}
  %
  Top: Electron density maps for the simulations using four different force fields.
  %
  Bottom left: Electron density map for the experiment. The scaling is uniform between all simulations and experiment. The original electron density profiles are shown in Fig.~\ref{SIfig:densprofs}. The effect of system size on the density profiles in simulations is demonstrated in Fig.~\ref{SIfig:densprofssize} in the SI.
  %
  Bottom center: Bilayer thickness. Thickness is defined as twice the distance from the peak in electron density to the membrane core. Experimental data are extracted in a similar manner from electron density profiles obtained with X-ray scattering.
  %
  Bottom right: area per lipid measured by dividing the total membrane area by the number of phospholipids. The experimental estimates, interpolated to 298~K, are taken from Refs.~\citenum{pabst2000structural} (58.8~\AA{}$^2$), \citenum{kuvcerka2006structure} (67.5~\AA{}$^2$), and \citenum{Kucerka11} (63.5~\AA{}$^2$). The NMR data is measured in Ref.~\citenum{Leftin14} using either $^{13}$C (58.9~\AA{}$^2$) or $^2$H (58.3~\AA{}$^2$) NMR.
  %
  The size-dependency of area per lipid is shown in Fig.~\ref{SIfig:aplvssize}.
  }
\end{figure}

Overall, all electron density profiles share the same features across all CHOL concentrations; a high-density band corresponding to the tightly-packed interfacial region containing electron-rich phoshorus, a low-density region at the core of the membrane occupied by the disordered acyl chains, and intermediate density in the rest of the lipid regions as well as the aqueous phase. However, a more detailed look at the profiles in Fig.~\ref{fig:densmaps} reveals differences between the force fields. CHOL renders the membrane thicker, resulting in the spreading of the high-density bands with increasing amount of CHOL. In CHARMM36, Slipids, and MacRog, there is clear saturation of this effect at around 30~mol-\% CHOL. In Lipid17, however, the membrane gets thicker until the maximum studied CHOL concentration of 47~mol-\%. In experiment, the effect of CHOL is generally weaker, and some saturation is evident at high CHOL concentrations. These trends are visible in the bottom middle panel in Fig.~\ref{fig:densmaps}, which shows the effect of CHOL on membrane thickness extracted from the density peak positions. Fig.~\ref{fig:densmaps} also demonstrates that CHARMM36 has the sharpest low- and high-density bands indicating smaller membrane fluctuations that would smear the bands, and the same is true for MacRog at higher cholesterol concentrations. The less sharp bands for Lipid17 and especially Slipids profiles indicate that the membranes are more flexible with these force fields.

Comparison against experiments (bottom left panel in Fig.~\ref{fig:densmaps}) indicates that the membrane structure is more uniform in the experiment, as the bands are sharper. This is also evident in the original electron density profiles in Fig.~\ref{SIfig:densprofs}. System size plays a role here, as demonstrated by the density maps calculated for the CHARMM36 simulations in three sizes, shown in Fig.~\ref{SIfig:densprofssize}. The larger the system, the larger the bilayer undulations are, which can smear the features of the density profiles. Still, even in the smallest system, the band intensities are not as localized as in the experiment. This difference likely partially stems from the features of the used SDP model, but it could also indicate different elastic properties between simulation and experiment.

@@

The membrane-condensing effect of CHOL is well documented, yet there is some ambiguity surrounding the term \cite{edholm2005areas}. The cross-sectional area of a CHOL molecule is smaller than that of a phospholipid, so an increase in CHOL concentration is expected to lower the average \emph{area per lipid} (including CHOL). However, condensation could also refer to a decrease in the \emph{area per phospholipid} upon the addition of cholesterol \cite{edholm2005areas}, or to CHOL having a diminishing partial area, meaning that a certain amount of CHOL could be added to a phospholipid bilayer without affecting its total area \cite{javanainen2017two}. We thus compared the behavior of the area per (POPC) lipid (APL) as a function of CHOL concentration in the bottom right panel of Fig.~\ref{fig:densmaps}. The dependence of APL on CHOL concentration follows the trends in thickness inversely in general, yet provide curious differences between the force fields. Lipid17 has the largest APL across the entire CHOL contration range, and the addition of CHOL always leads to an increase in the membrane area. MacRog also has a large APL for pure POPC, but the partial area of CHOL is negative until 30~mol-\% concentration indicating a particularly strong condensation effect. The profiles for Slipids and CHARMM36 are very similar, and the partial area of CHOL is small or even zero until a concentration of 20~mol-\%. Importantly, the observed differences in CHOL condensing effect take place at the physiologically relevant CHOL concentration range \cite{van2008membrane}. Experimental data are derived from models, so they can vary greatly, as evidenced by the very different values obtained for POPC by scattering and NMR experiments performed by different teams (See Fig.~\ref{fig:densmaps} and its caption). Still, all the simulation force fields provide areas per lipid that fall within the experimental spread for pure POPC.

@@

Order parameters for C--H bond vectors measured with with $^{13}$C or $^{2}$H NMR techniques, provide indirect information about the structural sampling of individual molecules \cite{ollila16}. Moreover, these parameters are readily extracted from the simulation. An increase in acyl chain order parameters with added cholesterol in simulations and experiments can be explained by am increased fraction of \textit{anti} conformations in the acyl chains \cite{ferreira13,??}, which is suggested to play critical role in the phase behaviour of PC--cholesterol mixtures \cite{ipsen87}. Consequently, the correct cholesterol ordering effect is expected to be a necessary condition for a model used to understand lipid--cholesterol phase behaviour.

Here, as the NMR experiments \cite{ferreira13} were again performed at different CHOL concentrations than simulations (and X-ray scattering measurements), we generated order parameter maps as a function of acyl chain carbon number and cholesterol concentration for both simulations and experiments, and subtracted these maps to obtain the deviation of simulation from experiment. These maps are shown for the palmitate (top row) and oleate (bottom row) chains of POPC in Fig.~\ref{fig:OPmaps}, whereas the original order parameter profiles are shown in Figs.~\ref{SIfig:palmitate} and \ref{fig:SIfig:oleate}.

\begin{figure}[htb!]
  \centering
  \includegraphics[width=\linewidth]{../FIGS/OP_chains.pdf}
  \caption{\label{fig:OPmaps}%
  POPC acyl chain order parameter deviation from experiments. Data are shown for palmitate (top row) and oleate (bottom row) and for the four force fields (columns). Negative values indicate that order is too high ($-S_\mathrm{CD}$ values too negative) in the simulations. The values that are within the estimated error range of 0.02 from experiment are considered to be ``correct'' and coloured in white. The order parameters calculated for the two or three identical carbons were very similar and thus averaged over in simulations, as they have identical values in experiments. For the C2 carbon of the oleate chain, the order parameters are forked in simulations and experiments. We thus calculated the differences for both the larger and smaller values, and averaged this deviation for the C2 column of oleate.
  }
\end{figure}

These original maps in Figs.~\ref{SIfig:palmitate} and \ref{fig:SIfig:oleate} demonstrate that the addition of cholesterol substantially increases the acyl chain C--H order parameters for all simulation models and experiments, which can be explained by increased \textit{anti} conformations in acyl chains \cite{ferreira13,??}. 
%The absolute values of order parameters of \textit{sn}-1 chain from $^{13}$C NMR experiments \cite{ferreira13} exhibit approximately linear increase with cholesterol up to equimolar mixture (Fig. \ref{slopes}), which is expected because phase separation is not observed for this mixture \cite{ionova12,ferreira13}.
The deviations mapped in Fig.~\ref{fig:OPmaps} provide an intuitive way for a quantitative comparison of the behaviors of the force fields. Here, deviations in the range [--0.02,0.02] are considered to fall within the error estimate and are thus shown in white. Blue indicates that the order parameters are too low in the simulation (their absolute value is larger and the chains are too ordered). Overall, the simulation models behave reasonably well at low CHOL concentrations, but deviate significantly from experiment at higher CHOL concentrations.

@@

The head group order parameters are not largely affected by CHOL concentration in experiments \cite{ferreira13}, which has been discussed in our earlier work \cite{botan15}. The four force fields evaluated here also reproduce this behavior to a great degree (Fig.~\ref{SIfig:headgroups}), so the limitations observed in the absence of CHOL \cite{botan15} largely persist across all studied CHOL concentrations. Moreover, it has now become standard to consider them in the parameterization and validation of lipid force fields \cite{dickson2022lipid21,grote2020optimization,yu2021charmm36,yu2021semi}.

\todo{Cholesterol order parameters? Into the deviation metric?}

@@

The previous sections highlight that the agreement of the studied force fields with experiment depends on both the parameter in question or the CHOL concentration. To capture these features in an easily digestible form, we have calculated the relative deviations of the form factor minima, the order parameters of the two acyl chains, and the diffusion coefficients between the experiments and all the force fields. The interpolated data were used, and the relative deviation from experiments was obtained by dividing the difference of simulated and experimental data by the latter. These data are shown in the two leftmost columns in Fig.~\ref{fig:deviation} as stacked plots.
